---
title: 'Part 2: CLM Practice (Homework 12)'
subtitle: 'Lab 2: What Makes a Product Successful? - W203 Section 8'
author: 'Team Herkimer: Rick Chen, Chase Madson, Maria Manna, Jash Sompalli'
date: '`r format(Sys.time(), "%b %d, %Y")`'
knit: (function(inputFile, encoding) { 
      out_dir <- '../reports';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), 
                        out_dir, 'homework_12.pdf')) })
output:
  pdf_document:
    number_sections: true
    toc: true
---

```{r Setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

```


```{r}
videos <- 
  '../data/external/videos.txt' %>% 
  read_tsv(col_types = 'cciciidii')

```


* views: the number of views by YouTube users.
* rate: This is the average of the ratings that the video received. You may think of this as a proxy for
video quality. (Notice that this is different from the variable ratings which is a count of the total
number of ratings that a video has received.)
* length: the duration of the video in seconds.

$ln(views)=\beta_0+\beta_1rate+\beta_2length$

```{r}
fit <-
  lm(data = videos, 
     formula = log(views) ~ rate + length)

summary(fit)

```


# Evaluate the IID assumption.


# No Perfect Colinearity

```{r, warning=FALSE}
videos %>% 
  ggplot(mapping = aes(x = rate, y = length)) + 
  geom_point()

```

# Linear Conditional Expectation


# Homoskedastic Errors

```{r}
plot(fit, 1)
```



# Normally Distributed Errors

```{r}
plot(fit, 2)

```

